Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-14 Manuscript/manuscript.tex
Calibration of lipid electrometer: Response of POPC head groups to bound charge

Before proceeding to the sodium and calcium  ion binding affinities, we quantify
the response of the headgroup order parameters to the amount of 
bound charge by using mixtures of monovalent cationic
surfactants (dihexadecyldimethylammonium)
and POPC~\cite{scherer89}. The amount of bound charge per PC 
in these systems is given simply by the molar fraction of cationic 
surfactants, because essentially all the surfactants with two hydrophopic
acyl chains can be assumed to locate in the lipid bilayers.
The experimental data for these systems can be used to validate 
the sensitivity of lipid headgroup order parameters
to the amount of bound charge in simulations.

The changes of the headgroup order parameters with increasing amount of 
the cationic surfactant are compared for simulations and experiments~\cite{scherer89} in Fig. \ref{OrderParameterCHANGESsurf}.
An approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ},
is observed in both simulations and experiments at least for mole fractions
below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that 
the head group order parameters respond too sensitively to the bound positive charge.
In contrast, the slope for the ECC-lipids model is in a very good agreement with experiments
for the $\alpha$ segment, while being slightly
underestimated for the $\beta$ segment.

The headgroup P-N vector angle with respect to the membrane normal
is also shown in Fig.~\ref{OrderParameterCHANGESsurf}
as a function of the mole fraction of the cationic surfactant.
As suggested previously~\cite{seelig87}, the headgroup orients
more towards the water phase with the increasing amount of positive
charge in a PC lipid bilayer. The effect is more pronounced in the
Lipid14 model, for which the addition of 50\% mole fraction of the cationic
surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle,
while the corresponding change in the ECC-lipids model is 11$^{\circ}$.
The difference is in line with the smaller order parameter 
changes and the reduced charge--dipole interactions in the latter model.
The lower sensitivity of the P-N vector angle response in the ECC-lipid
model can be considered to be more realistic, because the changes of the headgroup
order parameters as a function of the bound positive charge
are in better agreement with experiments in this model.
The results also imply that validation and improvements of MD simulation
models are generally required for reliable simulation studies of the
lipid headgroup responses to ions or other biomolecules, as also suggested
previously \cite{botan15}.

>>>>

Before studying the sodium and calcium ion binding affinities, we have to quantify as reference the response of the headgroup order parameters to the amount of bound charge by using mixtures of monovalent cationic surfactants (dihexadecyldimethylammonium) and POPC~\cite{scherer89}. These mixtures have a well-defined amount of bound charge per PC, i.e., the molar fraction of cationic surfactants. This direct relation results from the ability of dihexadecyldimethylammonium to be inserted in the lipid bilayers as any lipid because of its two hydrophobic acyl chains. Furthermore, available experimental data for these systems can be used to validate the sensitivity of lipid headgroup order parameters to the amount of bound charge in simulations \cite{REF}.

The changes of the headgroup order parameters with increasing amount of the cationic surfactant are compared between simulations and experiments~\cite{scherer89} in Fig.~\ref{OrderParameterCHANGESsurf}. An approximately linear decrease of the order parameters, as expected from Eq. \ref{OPchangeEQ}, is observed in both simulations and experiments at least for mole fractions below $\sim$30\%. The slope is, however, too steep in the Lipid14 model indicating that the head group order parameters response is too large to the bound positive charge. In contrast, the slope of the ECC-lipids model is in a very good agreement with experiments for the $\alpha$ segment, while being slightly underestimated for the $\beta$ segment.

In Fig.~\ref{OrderParameterCHANGESsurf}, we show the headgroup P-N vector angle as a function of the mole fraction of the cationic surfactant. As suggested previously~\cite{seelig87}, the headgroup orients more towards the water phase with the increasing amount of positive charge in a PC lipid bilayer. The effect is more pronounced in the Lipid14 model than in the ECC-POPC model.   For example, the addition of 50\% mole fraction of the cationic surfactant leads to the decrease of 20$^{\circ}$ of the P-N vector angle for the Lipid14 model while only 11$^{\circ}$ in the ECC-POPC model. The difference is in line with the smaller order parameter changes and the reduced charge--dipole interactions in the latter model. The lower sensitivity of the P-N vector angle response in the ECC-POPC model agrees better with experiments. The results imply that improvements are required in MD models for reliable studies of the lipid headgroup responses to ions or other biomolecules{\color{red}, as also suggested previously \cite{botan15}} \todo{Do you want to finish each paragraph with a citation to your previous works?}.
